Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2012-10-19 Dail_Mad_ext.tex
more simulation results

Generally, estimator bias and error were low with the exponential model, and 
coverage was nominal (Fig. ~\ref{fig:exp_hists}; Appendix ?). However, estimates of $r$ were 
imprecise and somewhat biased when the true value of $\Lambda$ was low (1). 
Simulations of the Ricker-logistic model also generally performed well (Figure 2; 
Appendix ?), except when the true value of $r$ was low (0.005). In this case the 
estimates of $K$ were imprecise and biased, especially when the true value of $K$ 
was high.  All three cases simulated for the Ricker-logistic + immigration model 
performed well (Figure 3; Appendix ?).

@@

Estimates for the top ranked model for ovenbird
(NB[$\Lambda (.) \alpha (.)$]Ricker-logistic+Immigration[$r(.)K(.) \iota
(.)$]p(wind+1st)) were similar when
run in the Bayesian framework, except that the estimate of r
more than halved (from 0.026 $\pm$ 0.006 to 0.011 $\pm$ 0.005 [for
Bayesian model estimates we present mean and SD]). When random observer effects
were added, estimates for $\Lambda$ and K increased dramatically
(from 31.6 $\pm$ 4.9 to 42.5 $\pm$ 7.3 and from 56.2 $\pm$ 18.2 to 101.8 $\pm$ 21.1,
respectively), and the estimate for r was intermediate (0.018 $\pm$ 0.006). The
estimate of the intercept for p (on the logit scale) dropped from -1.5 $\pm$
0.1 to -2.0 $\pm$ 0.1, and the estimate of $\sigma_p$ was 0.35 $\pm$ 0.03.

Gelman and Rubin diagnostics and visual examination of the chain trajectory and density plots
provided some evidence for a lack of convergence in estimates of the
intercept for p for all Bayesian models ran.

@@

Estimates for the top ranked model for golden-winged warbler
(ZIP[Λ(.)α(.)]Exponential[r(.)]p(1st))
were similar when run in the Bayesian framework. When random
observer effects were added, the estimate for $\Lambda$ increased dramatically
(from 22.4 $\pm$ 7.6 to 55.5 $\pm$ 30.9). The estimate of the intercept
for p (on the logit scale) dropped from -3.5 $\pm$ 0.2 to -5.6 $\pm$ 1.1, and the
estimate of σ was 1.7 $\pm$ 0.7. Adding zero-inflation to the dynamics (as well as
the initial abundance) has no effect for exponential models, because populations
that started at 0 would stay at 0 in any case. Therefore, we tested
zero-inflated dynamics on the Exponential + Immigration model. Adding zero-inflation to the dynamics
increased the estimate of ι from 0 to 0.091 $\pm$ 0.049, but had only minor
effects on the other parameters. Gelman and Rubin diagnostics and visual
examination of the chain trajectory and density plots provided evidence for a lack of convergence in
estimates of $\Lambda$ for all Bayesian models ran, and for most parameters when random
effects were included.

@@

>>>>

Generally, estimator bias and error were low with the exponential model, and 
coverage was nominal (93.4\% - 97.0\%; Fig.~\ref{fig:exp_hists}; Appendix ?). 
However, estimates of $r$ were 
imprecise (RMSE between 0.020 and 0.025; relative RMSE between 413\% and 
482\% for non-zero true $r$) and biased low (between -0.008 and -0.005; relative bias between
-151\% and -101\% for non-zero true $r$) when the true value of $\Lambda$ was low (1). 
Simulations of the Ricker-logistic model also generally performed well (Fig.~\ref{fig:rick_hists}; 
Appendix ?), except when the true value of $r$ was low (0.005). In this case the 
estimates of $r$ and $K$ were inaccurate (relative RMSE between 87.4\% and 2091.8\%),
biased high (between 65.4\% and 198.9\%), and had low coverage (between 77.6\% and 94.3\%).
In other cases for the Ricker-logistic model, relative RMSE was between 6.7\%
and 55\%, bias was between -6.4\% and 5.8\%, and coverage was between
91.9\% and 96.4\%.  All three cases simulated for the Ricker-logistic + immigration model 
performed fairly well (Fig.~\ref{fig:ricki_hists}; Appendix ?; relative RMSE between 7.2\% 
and 62.9\%, bias between -3.5\% and 3.3\%, and coverage between
93.2\% and 96.4\%).

@@

Estimates for the top ranked model for ovenbird
%(NB[$\Lambda (.) \alpha (.)$]Ricker-logistic+Immigration[$r(.)K(.) \iota(.)$]p(wind+1st)) 
were similar when
run in the Bayesian framework, except that the estimate of r
more than halved (from 0.026 $\pm$ 0.006 to 0.011 $\pm$ 0.005 [for
Bayesian model estimates we present mean and SD]). When random observer effects
were added, estimates for $\Lambda$ and K increased dramatically
(from 31.6 $\pm$ 4.9 to 42.5 $\pm$ 7.3 and from 56.2 $\pm$ 18.2 to 101.8 $\pm$ 21.1,
respectively), and the estimate for r was intermediate (0.018 $\pm$ 0.006). The
estimate of the intercept for p (on the logit scale) dropped from -1.5 $\pm$
0.1 to -2.0 $\pm$ 0.1, and the estimate of $\sigma_p$ was 0.35 $\pm$ 0.03.
Gelman and Rubin diagnostics and visual examination of the chain trajectory and density plots
provided some evidence for a lack of convergence in estimates of the
intercept for p for all Bayesian models ran.

@@

Estimates for the top ranked model for golden-winged warbler
%(ZIP[Λ(.)α(.)]Exponential[r(.)]p(1st))
were similar when run in the Bayesian framework. When random
observer effects were added, the estimate for $\Lambda$ increased dramatically
(from 22.4 $\pm$ 7.6 to 55.5 $\pm$ 30.9). The estimate of the intercept
for $p$ (on the logit scale) dropped from -3.5 $\pm$ 0.2 to -5.6 $\pm$ 1.1, and the
estimate of $\sigma_p$ was 1.7 $\pm$ 0.7. Adding zero-inflation to the dynamics (as well as
the initial abundance) has no effect for exponential models, because populations
that started at 0 would stay at 0 in any case. Therefore, we tested
zero-inflated dynamics on the Exponential + Immigration model. Adding zero-inflation to the dynamics
increased the estimate of ι from 0 to 0.091 $\pm$ 0.049, but had only minor
effects on the other parameters. Gelman and Rubin diagnostics and visual
examination of the chain trajectory and density plots provided evidence for a lack of convergence in
estimates of $\Lambda$ for all Bayesian models ran, and for most parameters when random
effects were included.

@@

\begin{sidewaysfigure}
  \centering
  \includegraphics{figs/rick_hists}
\caption{Histograms of 1000 parameter estimates for each of 9
simulation cases with Ricker-logistic dynamics. The vertical lines are the 
data-generating values.}
\label{fig:rick_hists}
\end{sidewaysfigure}

\begin{sidewaysfigure}
  \centering
  \includegraphics[width=8.5in]{figs/ricki_hists}
\caption{Histograms of 1000 parameter estimates for each of 3
simulation cases with Ricker-logistic + immigration dynamics. The vertical lines are the 
data-generating values.}
\label{fig:ricki_hists}
\end{sidewaysfigure}
